Channel Shaping Using Reconfigurable Intelligent Surfaces: From Diagonal to Beyond

### 2024-02-01 main.tex
:dizzy: Confused at proof

	\emph{Notation:}
	Italic, bold lower-case, and bold upper-case letters indicate scalars, vectors and matrices, respectively.
	$\jmath$ denotes the imaginary unit.
	$\mathbb{C}$ represents the set of complex numbers.
	% $\mathrm{U}(n)$ is the group of $n \times n$ unitary matrices.
	$\mathbb{U}^{n \times n}$ denotes the set of $n \times n$ unitary matrices.
	$\mathbf{0}$ and $\mathbf{I}$ are the all-zero and identity matrices with appropriate size, respectively.
	$\mathrm{tr}(\cdot)$ and $\det(\cdot)$ evaluates the trace and determinant of a square matrix, respectively.
	$\mathrm{diag}(\cdot)$ constructs a square matrix with arguments on the main diagonal and zeros elsewhere.
	$\mathrm{sv}(\cdot)$ returns the singular value vector.
	$\sigma_n(\cdot)$ and $\lambda_n(\cdot)$ is the $n$-th largest singular value and eigenvalue, respectively.
	% $\boldsymbol{\sigma}(\cdot)$ and $\boldsymbol{\lambda}(\cdot)$ are the corresponding vectors.
	$(\cdot)^*$, $(\cdot)^\mathsf{T}$, $(\cdot)^\mathsf{H}$, $(\cdot)^{(r)}$, $(\cdot)^{\star}$ denote the conjugate, transpose, conjugate transpose (Hermitian), $r$-th iterated point, and final solution, respectively.
	$\lvert \cdot \rvert$ denotes the absolute value.
	$\lVert \cdot \rVert _p$ means the $p$-norm and $\lVert \cdot \rVert$ suggests $p = 2$.
	$\lVert \cdot \rVert _\mathrm{F}$ represents the Frobenius norm.
	% $(\cdot)^{(r)}$ and $(\cdot)^{\star}$ denote the $r$-th iterated and optimal results, respectively.
	% The distribution of a \gls{cscg} random variable with zero mean and variance $\sigma^2$ is denoted by $\mathcal{CN}(0,\sigma^2)$, and $\sim$ means ``distributed as''.
	$\sim$ means ``distributed as''.
	$\mathcal{CN}(\mathbf{0}, \mathbf{\Sigma})$ is the multivariate \gls{cscg} distribution with mean $\mathbf{0}$ and covariance $\mathbf{\Sigma}$.
	% The \gls{cscg} distribution with zero mean and variance $\sigma^2$ is denoted by $\mathcal{CN}(0,\sigma^2)$.
\end{section}

\begin{section}{\gls{bd}-\gls{ris} Model}
	Consider a \gls{bd}-\gls{ris} aided point-to-point \gls{mimo} system with $N_\mathrm{T}$, $N_\mathrm{S}$, $N_\mathrm{R}$ transmit, scatter, and receive antennas, respectively.
	This configuration is denoted as $N_\mathrm{T} \times N_\mathrm{S} \times N_\mathrm{R}$.
	The \gls{bd}-\gls{ris} is modeled as an $N_\mathrm{S}$-port network \cite{Ivrlac2010} that further divides into $G$ individual groups.
	Each group contains $L \triangleq N_\mathrm{S} / G$ elements interconnected by real-time reconfigurable components \cite{Shen2020a}.
	% To simplify the analysis, we assume the \gls{bd}-\gls{ris} features lossless asymmetric\footnote{While symmetric impedance network is often considered in the literature \cite{Shen2020a,Nerini2023,Santamaria2023,Fang2023,Nerini2023a,Zhou2023,Li2023d,Bartoli2023}, asymmetric passive components (e.g., ring hybrids and branch-line hybrids) may also be reconfigured in real time \cite{Ahn2006}. Asymmetric \gls{bd}-\gls{ris} has been discussed in \cite{Li2023b,Li2023c,Bartoli2023}.} in-group connections without mutual coupling.
	To simplify the analysis, we assume there are no mutual coupling and the in-group connections can be lossless and asymmetric\footnote{While symmetric impedance network is often considered in the literature \cite{Shen2020a,Nerini2023,Santamaria2023,Fang2023,Nerini2023a,Zhou2023,Li2023d,Bartoli2023}, asymmetric passive components (e.g., ring hybrids and branch-line hybrids) may also be reconfigured in real time \cite{Ahn2006}. Asymmetric \gls{bd}-\gls{ris} has been discussed in \cite{Li2023b,Li2023c,Bartoli2023}.}.
	The overall scattering matrix is thus block diagonal $\mathbf{\Theta} = \mathrm{diag}(\mathbf{\Theta}_1,\ldots,\mathbf{\Theta}_G) \in \mathbb{U}^{N_\mathrm{S} \times N_\mathrm{S}}$, where $\mathbf{\Theta}_g \in \mathbb{U}^{L \times L}$ is a unitary matrix corresponding to group $g \in \mathcal{G} \triangleq \{1,\ldots,G\}$.
	Let $\mathbf{H}_\mathrm{D} \in \mathbb{C}^{N_\mathrm{R} \times N_\mathrm{T}}$, $\mathbf{H}_\mathrm{F} \in \mathbb{C}^{N_\mathrm{S} \times N_\mathrm{T}}$, $\mathbf{H}_\mathrm{B} \in \mathbb{C}^{N_\mathrm{R} \times N_\mathrm{S}}$ denote the direct (transmitter-receiver), forward (transmitter-\gls{ris}), and backward (\gls{ris}-receiver) channels, respectively.
	The equivalent channel is
	\begin{equation}
		\mathbf{H} = \mathbf{H}_\mathrm{D} + \mathbf{H}_\mathrm{B} \mathbf{\Theta} \mathbf{H}_\mathrm{F} = \mathbf{H}_\mathrm{D} + \sum_g \mathbf{H}_{\mathrm{B},g} \mathbf{\Theta}_g \mathbf{H}_{\mathrm{F},g},
		\label{eq:channel_equivalent}
	\end{equation}
	where $\mathbf{H}_{\mathrm{B},g} \in \mathbb{C}^{N_\mathrm{R} \times L}$ and $\mathbf{H}_{\mathrm{F},g} \in \mathbb{C}^{L \times N_\mathrm{T}}$ are the backward and forward channels of \gls{ris} group $g$, respectively.
	\begin{remark}
		\gls{bd}-\gls{ris} reduces to diagonal \gls{ris} and unitary \gls{ris} with group size 1 and $N_\mathrm{S}$, respectively.
	\end{remark}
	\begin{remark}
		Individual forward and backward \gls{csi} are required for \gls{bd}-\gls{ris} designs.
		This is different from diagonal \gls{ris} where estimating their product is usually sufficient.
		% Later we will show the potential benefits from the \gls{csi} overhead.
	\end{remark}
\end{section}

\begin{section}{Channel Singular Value Redistribution}
	\begin{subsection}{A Toy Example}
		We first illustrate the channel shaping capabilities of different \gls{ris} by a toy example.
		% We first use a toy example to illustrate that \gls{bd}-\gls{ris} can provide a wider dynamic range of channel singular values.
		Consider a $2 \times 2 \times 2$ setup where the direct link is blocked.
		The diagonal \gls{ris} is modeled by $\mathbf{\Theta}_\mathrm{D} = \mathrm{diag}(e^{\jmath \theta_1}, e^{\jmath \theta_2})$ while the unitary \gls{bd}-\gls{ris} has 4 independent angular parameters
		\begin{equation}
			\mathbf{\Theta}_\mathrm{U} = e^{\jmath \phi} \begin{bmatrix}
				e^{\jmath \alpha} \cos \psi  & e^{\jmath \beta} \sin \psi   \\
				-e^{-\jmath \beta} \sin \psi & e^{-\jmath \alpha} \cos \psi
			\end{bmatrix}.
		\end{equation}
		In particular, $\phi$ has no impact on the singular value because $\mathrm{sv}(e^{\jmath \phi} \mathbf{A}) = \mathrm{sv}(\mathbf{A})$.
		% For a fair comparison, we also consider symmetric impedance network with $\beta = \pi / 2$.
		We also enforce symmetry by $\beta = \pi / 2$ such that both architectures have the same number of angular parameters.
		% also enforce symmetry with $\beta = \pi / 2$.
		\begin{figure}
			\centering
			\includegraphics[width=\columnwidth]{assets/simulation/pc_singular_toy.eps}
			\caption{Channel singular value shaping by diagonal and symmetry unitary \gls{ris}. $(N^\mathrm{T}, N^\mathrm{S}, N^\mathrm{R}) = (2, 2, 2)$. Direct link is absent.}
			\label{sm:pc_singular_toy}
		\end{figure}
		Fig.~\ref{sm:pc_singular_toy} shows the channel singular values achieved by an exhaustive grid search over $(\theta_1, \theta_2)$ for diagonal \gls{ris} and $(\alpha, \psi)$ for symmetric unitary \gls{ris}.
		It is observed that both singular values can be manipulated up to $9\%$ using diagonal \gls{ris} and $42\%$ using symmetric \gls{bd}-\gls{ris}, despite both architectures have the same number of scattering elements and design parameters.
		A larger performance gap is expected when asymmetric \gls{bd}-\gls{ris} is available.
		This example shows \gls{bd}-\gls{ris} can provide a wider dynamic range of channel singular values and motivates further studies on channel shaping.
	\end{subsection}

	\begin{subsection}{Pareto Frontier Characterization}
		We then characterize the Pareto frontier of channel singular values by maximizing their weighted sum
		\begin{maxi!}
			{\scriptstyle{\mathbf{\Theta}}}{\sum_n \rho_n \sigma_n(\mathbf{H})}{\label{op:pareto}}{\label{ob:pareto}}
			\addConstraint{\mathbf{\Theta}_g^\mathsf{H} \mathbf{\Theta}_g=\mathbf{I},}{\quad \forall g,}{\label{co:pareto_unitary}}
		\end{maxi!}
		where $n \in \{1,\ldots,\min(N_\mathrm{T}, N_\mathrm{R})\}$ and $\rho_n$ is the weight of the $n$-th singular value that can be positive, zero, or negative.
		Varying $\{\rho_n\}$ unveils the entire achievable singular value region.
		Thus, the Pareto frontier problem~\eqref{op:pareto} generalizes most relevant metrics and provides a powerful shaping framework.
		% For group $g$, the objective \eqref{ob:pareto} is smooth in $\mathbf{\Theta}_g$ and the feasible domain \eqref{co:pareto.unitary} corresponds to the Stiefel manifold.
		The objective \eqref{ob:pareto} is smooth in $\mathbf{\Theta}$ and the feasible domain \eqref{co:pareto_unitary} for group $g$ corresponds to the Stiefel manifold.
		% Next, we zoom out to solve general smooth optimization problems of asymmetric \gls{bd}-\gls{ris} with arbitrary group size.
		% Those properties motivate us to propose a \gls{rcg} algorithm for smooth optimization problems of asymmetric \gls{bd}-\gls{ris} with arbitrary group size.
		% Those motivate us to propose a general \gls{rcg} algorithm for asymmetric \gls{bd}-\gls{ris} with arbitrary group size, which will be detailed in Section~\ref{sc:rcg}.
		% Those properties motivate us to zoom out to general smooth optimization problems of asymmetric \gls{bd}-\gls{ris} with arbitrary group size.
		Next, we zoom out to general smooth maximization problems of asymmetric \gls{bd}-\gls{ris} with arbitrary group size.

		Inspired by \cite{Abrudan2008,Abrudan2009}, we propose a block-wise \gls{rcg} algorithm along the geodesics on the Lie group of unitary matrices $\mathbb{U}^{L \times L}$.
		It leverages the fact that unitary matrices are closed under multiplication.
		At iteration $r$, the gradient is computed in the Euclidean space and translated to the Riemannian manifold \cite{Absil2009}
		\begin{gather}
			\nabla_{\mathrm{E},g}^{(r)} = \frac{\partial f(\mathbf{\Theta}_g^{(r)})}{\partial \mathbf{\Theta}_g^*},\label{eq:gradient_euclidean}\\
			\nabla_{\mathrm{R},g}^{(r)} = \nabla_{\mathrm{E},g}^{(r)} {\mathbf{\Theta}_g^{(r)}}^\mathsf{H} - \mathbf{\Theta}_g^{(r)} {\nabla_{\mathrm{E},g}^{(r)}}^\mathsf{H}.\label{eq:gradient_riemannian}
		\end{gather}
		The Polak-Ribierre parameter \cite{Polak1969} is approximated as \cite{Abrudan2009}
		\begin{equation}
			\gamma_g^{(r)} = \frac{\mathrm{tr}\bigl((\nabla_{\mathrm{R},g}^{(r)} - \nabla_{\mathrm{R},g}^{(r-1)}) {\nabla_{\mathrm{R},g}^{(r)}}^\mathsf{H}\bigr)}{\mathrm{tr}\bigl(\nabla_{\mathrm{R},g}^{(r-1)} {\nabla_{\mathrm{R},g}^{(r-1)}}^\mathsf{H}\bigr)},
			\label{eq:parameter_cg}
		\end{equation}
		and the conjugate direction is
		\begin{equation}
			\mathbf{D}_g^{(r)} = \nabla_{\mathrm{R},g}^{(r)} + \gamma_g^{(r)} \mathbf{D}_g^{(r-1)}.
			\label{eq:direction_cg}
		\end{equation}
		% In the Stiefel manifold, the geodesic emanating from $\mathbf{\Theta}_g^{(r)}$ with velocity $\mathbf{D}_g^{(r)}$ and step size $\mu_g$ is described compactly by the exponential map \cite{Edelman1998}
		% In the Stiefel manifold, the geodesic emanating from $\mathbf{\Theta}_g^{(r)}$ with velocity $\mathbf{D}_g^{(r)}$ and step size $\mu_g^{(r)}$ is described compactly by the exponential map \cite{Edelman1998}
		In the Stiefel manifold, the geodesic emanating from $\mathbf{\Theta}_g^{(r)}$ with velocity $\mathbf{D}_g^{(r)}$ and step size $\mu$ is described compactly by the exponential map \cite{Edelman1998}
		\begin{equation}
			\mathbf{G}_g^{(r)}(\mu) = \exp(\mu \mathbf{D}_g^{(r)}) \mathbf{\Theta}_g^{(r)}.
			\label{eq:geodesic}
		\end{equation}
		% \begin{gather}
		% 	\mathbf{T}_g^{(r)}(\mu) = \exp(\mu \mathbf{D}_g^{(r)}),\label{eq:exponential_map}\\
		% 	\mathbf{G}_g^{(r)}(\mu) = \mathbf{T}_g^{(r)}(\mu) \mathbf{\Theta}_g^{(r)}.\label{eq:geodesic}
		% \end{gather}
		An appropriate $\mu^\star$ can be obtained by the Armijo rule \cite{Armijo1966}.\footnote{To double the step size, one only need to square the rotation matrix instead of recomputing the matrix exponential, i.e., $\exp(2 \mu \mathbf{D}_g^{(r)}) = \exp^2(\mu \mathbf{D}_g^{(r)})$.}
		Finally, the scattering matrix is updated along the geodesic as
		\begin{equation}
			\mathbf{\Theta}_g^{(r+1)} = \mathbf{G}_g^{(r)}(\mu^\star).
			\label{eq:update_geodesic}
		\end{equation}
		% Algorithm~\ref{ag:rcg} summarizes the proposed block-wise geodesic \gls{rcg} update for asymmetric \gls{bd}-\gls{ris} with arbitrary group size.
		Algorithm~\ref{ag:rcg} summarizes the proposed block-wise geodesic \gls{rcg} method for general smooth maximization problems of asymmetric \gls{bd}-\gls{ris}.
		Convergence to stationary points is guaranteed as updating each block iteratively does not reduce the objective function.

		\begin{remark}
			Compared with universal manifold optimization \cite{Absil2009,Pan2022d}, Algorithm~\ref{ag:rcg} inherits a trifold benefit from \cite{Abrudan2008,Abrudan2009}:
			% The advantage of the proposed block-wise \gls{rcg} algorithm over universal manifold optimization \cite{Absil2009,Pan2022d} is trifold:
			\begin{enumerate}
				\item No retraction thanks to rotational update;
				\item Lower computational complexity per iteration;
				\item Faster convergence thanks to proper parameter space.
			\end{enumerate}
		\end{remark}

>>>>

	\emph{Notation:}
	Italic, bold lower-case, and bold upper-case letters indicate scalars, vectors and matrices, respectively.
	$\jmath$ denotes the imaginary unit.
	$\mathbb{C}$ represents the set of complex numbers.
	$\mathbb{U}^{n \times n}$ denotes the set of $n \times n$ unitary matrices.
	$\mathbf{0}$ and $\mathbf{I}$ are the all-zero and identity matrices with appropriate size, respectively.
	$\mathrm{tr}(\cdot)$ and $\det(\cdot)$ evaluates the trace and determinant of a square matrix, respectively.
	$\mathrm{diag}(\cdot)$ constructs a square matrix with arguments on the main diagonal and zeros elsewhere.
	$\mathrm{sv}(\cdot)$ returns the singular value vector.
	$\sigma_n(\cdot)$ and $\lambda_n(\cdot)$ is the $n$-th largest singular value and eigenvalue, respectively.
	% $\boldsymbol{\sigma}(\cdot)$ and $\boldsymbol{\lambda}(\cdot)$ are the corresponding vectors.
	$(\cdot)^*$, $(\cdot)^\mathsf{T}$, $(\cdot)^\mathsf{H}$, $(\cdot)^{(r)}$, $(\cdot)^{\star}$ denote the conjugate, transpose, conjugate transpose (Hermitian), $r$-th iterated point, and final solution, respectively.
	$\lvert \cdot \rvert$ denotes the absolute value.
	$\lVert \cdot \rVert _p$ means the $p$-norm and $\lVert \cdot \rVert$ suggests $p = 2$.
	$\lVert \cdot \rVert _\mathrm{F}$ represents the Frobenius norm.
	$\sim$ means ``distributed as''.
	$\mathcal{CN}(\mathbf{0}, \mathbf{\Sigma})$ is the multivariate \gls{cscg} distribution with mean $\mathbf{0}$ and covariance $\mathbf{\Sigma}$.
\end{section}

\begin{section}{\gls{bd}-\gls{ris} Model}
	Consider a \gls{bd}-\gls{ris} aided point-to-point \gls{mimo} system with $N_\mathrm{T}$, $N_\mathrm{S}$, $N_\mathrm{R}$ transmit, scatter, and receive antennas, respectively.
	This configuration is denoted as $N_\mathrm{T} \times N_\mathrm{S} \times N_\mathrm{R}$.
	The \gls{bd}-\gls{ris} is modeled as an $N_\mathrm{S}$-port network \cite{Ivrlac2010} that further divides into $G$ individual groups.
	Each group contains $L \triangleq N_\mathrm{S} / G$ elements interconnected by real-time reconfigurable components \cite{Shen2020a}.
	To simplify the analysis, we assume there are no mutual coupling and the in-group connections can be lossless and asymmetric\footnote{While symmetric impedance network is often considered in the literature \cite{Shen2020a,Nerini2023,Santamaria2023,Fang2023,Nerini2023a,Zhou2023,Li2023d,Bartoli2023}, asymmetric passive components (e.g., ring hybrids and branch-line hybrids) may also be reconfigured in real time \cite{Ahn2006}. Asymmetric \gls{bd}-\gls{ris} has been discussed in \cite{Li2023b,Li2023c,Bartoli2023}.}.
	The overall scattering matrix is thus block diagonal $\mathbf{\Theta} = \mathrm{diag}(\mathbf{\Theta}_1,\ldots,\mathbf{\Theta}_G) \in \mathbb{U}^{N_\mathrm{S} \times N_\mathrm{S}}$, where $\mathbf{\Theta}_g \in \mathbb{U}^{L \times L}$ is a unitary matrix corresponding to group $g \in \mathcal{G} \triangleq \{1,\ldots,G\}$.
	Let $\mathbf{H}_\mathrm{D} \in \mathbb{C}^{N_\mathrm{R} \times N_\mathrm{T}}$, $\mathbf{H}_\mathrm{F} \in \mathbb{C}^{N_\mathrm{S} \times N_\mathrm{T}}$, $\mathbf{H}_\mathrm{B} \in \mathbb{C}^{N_\mathrm{R} \times N_\mathrm{S}}$ denote the direct (transmitter-receiver), forward (transmitter-\gls{ris}), and backward (\gls{ris}-receiver) channels, respectively.
	The equivalent channel is
	\begin{equation}
		\mathbf{H} = \mathbf{H}_\mathrm{D} + \mathbf{H}_\mathrm{B} \mathbf{\Theta} \mathbf{H}_\mathrm{F} = \mathbf{H}_\mathrm{D} + \sum_g \mathbf{H}_{\mathrm{B},g} \mathbf{\Theta}_g \mathbf{H}_{\mathrm{F},g},
		\label{eq:channel_equivalent}
	\end{equation}
	where $\mathbf{H}_{\mathrm{B},g} \in \mathbb{C}^{N_\mathrm{R} \times L}$ and $\mathbf{H}_{\mathrm{F},g} \in \mathbb{C}^{L \times N_\mathrm{T}}$ are the backward and forward channels of \gls{ris} group $g$, respectively.
	\begin{remark}
		\gls{bd}-\gls{ris} reduces to diagonal \gls{ris} and unitary \gls{ris} with group size 1 and $N_\mathrm{S}$, respectively.
	\end{remark}
	\begin{remark}
		Individual forward and backward \gls{csi} are required for \gls{bd}-\gls{ris} designs.
		This is different from diagonal \gls{ris} where estimating their product is usually sufficient.
		% Later we will show the potential benefits from the \gls{csi} overhead.
	\end{remark}
\end{section}

\begin{section}{Channel Singular Value Redistribution}
	\begin{subsection}{A Toy Example}
		We first illustrate the channel shaping capabilities of different \gls{ris} by a toy example.
		% We first use a toy example to illustrate that \gls{bd}-\gls{ris} can provide a wider dynamic range of channel singular values.
		Consider a $2 \times 2 \times 2$ setup where the direct link is blocked.
		The diagonal \gls{ris} is modeled by $\mathbf{\Theta}_\mathrm{D} = \mathrm{diag}(e^{\jmath \theta_1}, e^{\jmath \theta_2})$ while the unitary \gls{bd}-\gls{ris} has 4 independent angular parameters
		\begin{equation}
			\mathbf{\Theta}_\mathrm{U} = e^{\jmath \phi} \begin{bmatrix}
				e^{\jmath \alpha} \cos \psi  & e^{\jmath \beta} \sin \psi   \\
				-e^{-\jmath \beta} \sin \psi & e^{-\jmath \alpha} \cos \psi
			\end{bmatrix}.
		\end{equation}
		In particular, $\phi$ has no impact on the singular value because $\mathrm{sv}(e^{\jmath \phi} \mathbf{A}) = \mathrm{sv}(\mathbf{A})$.
		We also enforce symmetry by $\beta = \pi / 2$ such that both architectures have the same number of angular parameters.
		\begin{figure}
			\centering
			\includegraphics[width=\columnwidth]{assets/simulation/pc_singular_toy.eps}
			\caption{$2 \times 2 \times 2$ channel singular value shaping by diagonal and symmetry unitary \gls{ris}. Direct link is absent.}
			\label{sm:pc_singular_toy}
		\end{figure}
		Fig.~\ref{sm:pc_singular_toy} shows the channel singular values achieved by an exhaustive grid search over $(\theta_1, \theta_2)$ for diagonal \gls{ris} and $(\alpha, \psi)$ for symmetric unitary \gls{ris}.
		It is observed that both singular values can be manipulated up to $9\%$ using diagonal \gls{ris} and $42\%$ using symmetric \gls{bd}-\gls{ris}, despite both architectures have the same number of scattering elements and design parameters.
		A larger performance gap is expected when asymmetric \gls{bd}-\gls{ris} is available.
		This example shows \gls{bd}-\gls{ris} can provide a wider dynamic range of channel singular values and motivates further studies on channel shaping.
	\end{subsection}

	\begin{subsection}{Pareto Frontier Characterization}
		We then characterize the Pareto frontier of channel singular values by maximizing their weighted sum
		\begin{maxi!}
			{\scriptstyle{\mathbf{\Theta}}}{\sum_n \rho_n \sigma_n(\mathbf{H})}{\label{op:pareto}}{\label{ob:pareto}}
			\addConstraint{\mathbf{\Theta}_g^\mathsf{H} \mathbf{\Theta}_g=\mathbf{I},}{\quad \forall g,}{\label{co:pareto_unitary}}
		\end{maxi!}
		where $n \in \{1,\ldots,\min(N_\mathrm{T}, N_\mathrm{R})\}$ and $\rho_n$ is the weight of the $n$-th singular value that can be positive, zero, or negative.
		Varying $\{\rho_n\}$ unveils the entire achievable singular value region.
		Thus, the Pareto frontier problem~\eqref{op:pareto} generalizes most relevant metrics and provides a powerful shaping framework.
		The objective \eqref{ob:pareto} is smooth in $\mathbf{\Theta}$ and the feasible domain \eqref{co:pareto_unitary} for group $g$ corresponds to the Stiefel manifold.
		Next, we zoom out to general smooth maximization problems of asymmetric \gls{bd}-\gls{ris}.

		Inspired by \cite{Abrudan2008,Abrudan2009}, we propose a block-wise \gls{rcg} algorithm along the geodesics on the Lie group of unitary matrices $\mathbb{U}^{L \times L}$.
		It leverages the fact that unitary matrices are closed under multiplication.
		At iteration $r$, the gradient is computed in the Euclidean space and translated to the Riemannian manifold \cite{Absil2009}
		\begin{gather}
			\nabla_{\mathrm{E},g}^{(r)} = \frac{\partial f(\mathbf{\Theta}_g^{(r)})}{\partial \mathbf{\Theta}_g^*},\label{eq:gradient_euclidean}\\
			\nabla_{\mathrm{R},g}^{(r)} = \nabla_{\mathrm{E},g}^{(r)} {\mathbf{\Theta}_g^{(r)}}^\mathsf{H} - \mathbf{\Theta}_g^{(r)} {\nabla_{\mathrm{E},g}^{(r)}}^\mathsf{H}.\label{eq:gradient_riemannian}
		\end{gather}
		The Polak-Ribierre parameter \cite{Polak1969} is approximated as \cite{Abrudan2009}
		\begin{equation}
			\gamma_g^{(r)} = \frac{\mathrm{tr}\bigl((\nabla_{\mathrm{R},g}^{(r)} - \nabla_{\mathrm{R},g}^{(r-1)}) {\nabla_{\mathrm{R},g}^{(r)}}^\mathsf{H}\bigr)}{\mathrm{tr}\bigl(\nabla_{\mathrm{R},g}^{(r-1)} {\nabla_{\mathrm{R},g}^{(r-1)}}^\mathsf{H}\bigr)},
			\label{eq:parameter_cg}
		\end{equation}
		and the conjugate direction is
		\begin{equation}
			\mathbf{D}_g^{(r)} = \nabla_{\mathrm{R},g}^{(r)} + \gamma_g^{(r)} \mathbf{D}_g^{(r-1)}.
			\label{eq:direction_cg}
		\end{equation}
		In the Stiefel manifold, the geodesic emanating from $\mathbf{\Theta}_g^{(r)}$ with velocity $\mathbf{D}_g^{(r)}$ and step size $\mu$ is described compactly by the exponential map \cite{Edelman1998}
		\begin{equation}
			\mathbf{G}_g^{(r)}(\mu) = \exp(\mu \mathbf{D}_g^{(r)}) \mathbf{\Theta}_g^{(r)}.
			\label{eq:geodesic}
		\end{equation}
		An appropriate $\mu^\star$ can be obtained by the Armijo rule \cite{Armijo1966}.\footnote{To double the step size, one only need to square the rotation matrix instead of recomputing the matrix exponential, i.e., $\exp(2 \mu \mathbf{D}_g^{(r)}) = \exp^2(\mu \mathbf{D}_g^{(r)})$.}
		Finally, the scattering matrix is updated along the geodesic as
		\begin{equation}
			\mathbf{\Theta}_g^{(r+1)} = \mathbf{G}_g^{(r)}(\mu^\star).
			\label{eq:update_geodesic}
		\end{equation}
		Algorithm~\ref{ag:rcg} summarizes the proposed block-wise geodesic \gls{rcg} method for smooth maximization problems of asymmetric \gls{bd}-\gls{ris}.
		Convergence to stationary points is guaranteed as updating each block iteratively does not reduce the objective function.

		\begin{remark}
			Compared with universal manifold optimization \cite{Absil2009,Pan2022d}, Algorithm~\ref{ag:rcg} inherits a trifold benefit from \cite{Abrudan2008,Abrudan2009}:
			\begin{enumerate}
				\item No retraction thanks to rotational update \eqref{eq:geodesic}, \eqref{eq:update_geodesic};
				\item Lower computational complexity per iteration;
				\item Faster convergence thanks to proper parameter space.
			\end{enumerate}
		\end{remark}

		\setalgorithmcaptionfont{\small}
		\begin{algorithm}[!t]
			\small
			\caption{Block-wise geodesic \gls{rcg} for asymmetric \gls{bd}-\gls{ris}}
			\label{ag:rcg}
			\begin{algorithmic}[1]
				\Require $f(\mathbf{\Theta})$, $G$
				\Ensure $\mathbf{\Theta}^\star$
				\Initialize {$r \gets 0$, $\mathbf{\Theta}^{(0)}$}
				\Repeat
					\For {$g \gets 1$ to $G$}
						\State $\nabla_{\mathrm{E},g}^{(r)} \gets$ \eqref{eq:gradient_euclidean} \label{ln:gradient_euclidean}
						\State $\nabla_{\mathrm{R},g}^{(r)} \gets$ \eqref{eq:gradient_riemannian}
						\State $\gamma_g^{(r)} \gets$ \eqref{eq:parameter_cg}
						\State $\mathbf{D}_g^{(r)} \gets$ \eqref{eq:direction_cg}
						\If {$\Re\bigl\{\mathrm{tr}({\mathbf{D}_g^{(r)}}^\mathsf{H} \nabla_{\mathrm{R},g}^{(r)})\bigr\} < 0$} \Comment{not an ascent direction}
							\State $\mathbf{D}_g^{(r)} \gets \nabla_{\mathrm{R},g}^{(r)}$
						\EndIf
						\State $\mu \gets 1$
						\State $\mathbf{G}_g^{(r)}(\mu) \gets$ \eqref{eq:geodesic}
						\While {$f\bigl(\mathbf{G}_g^{(r)}(2\mu)\bigr) - f(\mathbf{\Theta}_g^{(r)}) \ge \mu \cdot \mathrm{tr}(\mathbf{D}_g^{(r)} {\mathbf{D}_g^{(r)}}^\mathsf{H}) / 2$}
							\State $\mu \gets 2 \mu$
						\EndWhile
						\While {$f\bigl(\mathbf{G}_g^{(r)}(\mu)\bigr) - f(\mathbf{\Theta}_g^{(r)}) < \mu / 2 \cdot \mathrm{tr}(\mathbf{D}_g^{(r)} {\mathbf{D}_g^{(r)}}^\mathsf{H}) / 2$}
							\State $\mu \gets \mu / 2$
						\EndWhile
						\State $\mathbf{\Theta}_g^{(r+1)} \gets$ \eqref{eq:update_geodesic}
					\EndFor
					\State $r \gets r+1$
				\Until $\lvert f(\mathbf{\Theta}^{(r)}) - f(\mathbf{\Theta}^{(r-1)}) \rvert / f(\mathbf{\Theta}^{(r-1)}) \le \epsilon$
			\end{algorithmic}
		\end{algorithm}

		Back to the Pareto channel singular value problem \eqref{op:pareto}. The Euclidean gradient of \gls{bd}-\gls{ris} group $g$ is computed as
		\begin{equation}
			\frac{\partial \sum_n \rho_n \sigma_n(\mathbf{H})}{\partial \mathbf{\Theta}_g^*} = \mathbf{H}_{\mathrm{B},g}^\mathsf{H} \mathbf{U} \mathrm{diag}(\rho_1,\ldots,\rho_N) \mathbf{V}^\mathsf{H} \mathbf{H}_{\mathrm{F},g}^\mathsf{H},
			\label{eq:pareto_gradient}
		\end{equation}
		where $\mathbf{U} \in \mathbb{C}^{N_\mathrm{R} \times N}$ and $\mathbf{V} \in \mathbb{C}^{N_\mathrm{T} \times N}$ are the left and right compact singular matrices of $\mathbf{H}$, respectively.
		Algorithm~\ref{ag:rcg} can thus be invoked for the Pareto singular value problem \eqref{op:pareto}, where expression \eqref{eq:pareto_gradient} should be used in line \ref{ln:gradient_euclidean}.
	\end{subsection}

	\begin{subsection}{Some Analytical Bounds}
		% We then analyze the channel shaping \emph{capability} of \gls{bd}-\gls{ris} under specific setups.
		We then discuss some analytical bounds related to channel singular values.
			% \gls{bd}-\gls{ris} cannot achieve a higher \gls{dof} than diagonal \gls{ris}.
			\begin{proposition}
				In rank-deficient point-to-point channel, \gls{bd}-\gls{ris} cannot achieve a higher \gls{dof} than diagonal \gls{ris}.
			\end{proposition}
			\begin{proof}
				The scattering matrix of \gls{bd}-\gls{ris} can be decomposed as
				\begin{equation}
					\mathbf{\Theta} = \mathbf{L} \mathbf{\Theta}_\mathrm{D} \mathbf{R}^\mathsf{H},
				\end{equation}
				where $\mathbf{L}, \mathbf{R} \in \mathbb{U}^{N_\mathrm{S} \times N_\mathrm{S}}$ are block diagonal matrices of $L \times L$ unitary blocks and $\mathbf{\Theta}_\mathrm{D} \in \mathbb{U}^{N_\mathrm{S} \times N_\mathrm{S}}$ corresponds to diagonal \gls{ris}.
				% However, manipulating $\mathbf{L}$ and $\mathbf{R}$ does not change the rank of $\bar{\mathbf{H}}_\mathrm{B} \triangleq \mathbf{H}_\mathrm{B} \mathbf{L}$ and $\bar{\mathbf{H}}_\mathrm{F} \triangleq \mathbf{R}^\mathsf{H} \mathbf{H}_\mathrm{F}$.
				Manipulating $\mathbf{L}$ and $\mathbf{R}$ only rotates the linear spans of $\bar{\mathbf{H}}_\mathrm{B} \triangleq \mathbf{H}_\mathrm{B} \mathbf{L}$ and $\bar{\mathbf{H}}_\mathrm{F} \triangleq \mathbf{R}^\mathsf{H} \mathbf{H}_\mathrm{F}$ and keeps their rank.
				On the other hand, there exists
				\begin{equation}
					\arg_{\mathbf{\Theta}_\mathrm{D}} \max \ \mathrm{rank}(\mathbf{H}_\mathrm{B} \mathbf{\Theta}_\mathrm{D} \mathbf{H}_\mathrm{F}) = \min \bigl( \mathrm{rank}(\bar{\mathbf{H}}_\mathrm{B}), \mathrm{rank}(\mathbf{\Theta}_\mathrm{D}), \mathrm{rank}(\bar{\mathbf{H}}_\mathrm{F}) \bigr).
				\end{equation}
				% Therefore,
				% \begin{align}
				% 	\mathrm{rank}(\mathbf{H}_\mathrm{B} \mathbf{\Theta} \mathbf{H}_\mathrm{F}) & = \mathrm{rank}(\mathbf{H}_\mathrm{B} \mathbf{L} \mathbf{\Theta}_\mathrm{D} \mathbf{R}^\mathsf{H} \mathbf{H}_\mathrm{F}) \\
				% 	& = \mathrm{rank}(\bar{\mathbf{H}}_\mathrm{B} \mathbf{\Theta}_\mathrm{D} \bar{\mathbf{H}}_\mathrm{F}) \\
				% 	& \le \min \bigl( \mathrm{rank}(\bar{\mathbf{H}}_\mathrm{B}), \mathrm{rank}(\mathbf{\Theta}_\mathrm{D}), \mathrm{rank}(\bar{\mathbf{H}}_\mathrm{F}) \bigr).
				% \end{align}
				% On the other hand, we have
				% \begin{align}
				% 	\mathrm{rank}(\mathbf{H}_\mathrm{B} \mathbf{\Theta} \mathbf{H}_\mathrm{F})
				% \end{align}
				% Therefore,
				% \begin{align}
